---
title: "Shiny Appliation - mtcars"
output: ioslides_presentation
author: Jianlei (John) Sun
date: November 15, 2016
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Reproducible Pitch for mtcars

- This project consits of two parts. First, we need to create a Shiny application and deploy it on Rstudio's servers. Second, we should use the Rstudio Presenter to prepare a reproducible pitch presentation about the application. 
- Please find datasets, source code, and outputs from the following link.
 URL: https://github.com/jianleisun/data-science/tree/master/Shiny_Application_and_Reproducible_Pitch
- This presentation includes a brief introduction to the data input and source code.

## Dataset Information - mtcars

### Motor Trend Car Road Tests

The data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973–74 models).

### Reference
Henderson and Velleman (1981), Building multiple regression models interactively. Biometrics, 37, 391–411.

```{r}
library(datasets)
head(mtcars, 3)
```

## Dataset Nomenclature

**32 observations and 11 variables.**

| Index | Field | Detail |
------- | ----- | ------ |
| [, 1] | mpg | Miles/(US) gallon |
| [, 2]  | cyl | Number of cylinders |
| [, 3]	| disp | Displacement (cu.in.) |
| [, 4]	| hp | Gross horsepower |
| [, 5]	| drat | Rear axle ratio |
| [, 6]	| wt | Weight (lb/1000) |
| [, 7]	| qsec | 1/4 mile time |
| [, 8]	| vs | V/S |
| [, 9]	| am | Transmission (0 = automatic, 1 = manual) |
| [,10]	| gear | Number of forward gears |
| [,11]	| carb | Number of carburetors |

## Analysis - main code
```r
shinyServer(function(input, output) {
    
   ...
    
    fit <- reactive({
        lm(as.formula(txt1()), data=data)
    })
    
    ...
    
    output$fit <- renderPrint({
        summary(fit())
    })
    
    output$mpgPlot <- renderPlot({
        with(data, {
            plot(as.formula(txt1()))
            abline(fit(), col=2)
        })
    })
    
})
```


